---
title: "Simple document"
author: "Liliang Su"
date: "2025.9.6"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
```


## Importing data tables

Let's import a dataset.

```{r}
litters_df = read_csv(file = "./data/FAS_litters.csv")
#litters_df = read_csv(file = "~/Desktop/CU_course/8105_DataScience1/Topic2_DataWrangling_1/data/FAS_litters.csv")
```

Look at your data!

```{r}
names(litters_df)
```

Update the name in `litters_df`.
Convert column names to lower snake case. no space, replace it with underline.
```{r}
litters_df = 
  janitor::clean_names(litters_df)
```
Look at the data again

```{r}
litters_df
```

```{r}
tail(litters_df, 5)
```
```{r}
skimr::skim(litters_df)
```

*Note*: `view(litters_df)` only in the console, remove them from .Rmd

## Arguments to read_*

```{r}
litters_df = 
    read_csv(file = "./data/FAS_litters.csv",
    skip = 10, col_names = FALSE)
head(litters_df)


```

### Fix the missingness
```{r}
litters_df = 
    read_csv(
        file = "./data/FAS_litters.csv",
        na = c(".", "NA", ""))
head(litters_df)
```
learning assignment: importing `FAS_pups`

```{r}
pups_df =
  read_csv(
    "data/FAS_pups.csv",
    skip = 3,
    na = c( "NA",".", "")
    )

pups_df =
  janitor::clean_names(pups_df)
```

## Parsing columns
```{r}
litters_df = 
    read_csv(file = "./data/FAS_litters.csv",
        na = c(".", "NA", ""),
    col_types = cols(
      Group = col_character(),
      `Litter Number` = col_character(),
      `GD0 weight` = col_double(),
      `GD18 weight` = col_double(),
      `GD of Birth` = col_integer(),
      `Pups born alive` = col_integer(),
      `Pups dead @ birth` = col_integer(),
      `Pups survive` = col_integer()
    )
  )

tail(litters_df)
```

```{r}
litters_df = 
    read_csv(file = "./data/FAS_litters.csv",
        na = c(".", "NA", ""),
    col_types = cols(
      Group = col_factor()
    )
)

head(litters_df)
```

learning assignments
```{r}
pups_df = 
    read_csv("./data/FAS_pups.csv",
        na = c(".", "NA"), col_types = "fddddd")

skimr::skim(pups_df)
head(pups_df)
```

## xlsx
```{r}
library(readxl)

mlb11_df = read_excel("data/mlb11.xlsx", n_max = 20)

head(mlb11_df, 5)
```

```{r}
fellowship_ring = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6") 
```


## other file formats
```{r}
library(haven)

pulse_df = read_sas("./data/public_pulse_data.sas7bdat")

pulse_df = janitor::clean_names(pulse_df)
```

## Comparsion with base R (read.csv)
```{r}
pups_base = read.csv("./data/FAS_pups.csv")
pups_readr = read_csv("./data/FAS_pups.csv")

pups_base
pups_readr

pups_base$S
pups_readr$S
```